Tilting at Algorithmic Windmills

### 2013-05-14 Tilting/Rpaper.Rnw
WIP: reorganize slightly

\section{Guidelines}

\subsection{Include parameters.}
A best practice exemplar of comparative evaluation studies
is the work by Klein et al. \cite{Klein2009} in which
he compared 14 nonlinear registration algorithms using
four publicly available labeled brain data sets.  
As part of the study, the authors themselves tuned 
the parameters for their algorithm to ensure good
performance.  As
part of this work, Prof. Klein availed the scripts
used for the study on his website.%
\footnote{
http://mindboggle.info/papers/evaluation\_NeuroImage2009.php
}
This practice promotes reproducibility, fairness, and transparency in 
the evaluation and is perhaps most fundamental in these types
of studies for establishing conclusions regarding the relative
performance of algorithms.

@@

\subsection{Hassle the developers for a set of default parameters.}
Dovetailing with the previous item, developers should also facilitate
comparisons by providing default parameter files, sample scripts, 
or possible command line calls.  For example, the developers of
the image registration program Elastix \cite{klein2010} provide an
assortment of parameter files on the wiki page%
\footnote{
  http://elastix.bigr.nl/wiki/index.php/Parameter\_file\_database
}
listed in tabular format complete with  short
description (including applied modality and object of interest) and 
any publications which used that specific parameter file.  Additionally,
other packages are associated with discussion forums where users can
query regarding usage.

\subsection{Do not code up your own version of the algorithm particularly 
if one is available from the original authors.  If you do, be sure to make
your implementation available.}

\subsection{Provide details as to the source of the algorithm}
As a corollary to the previous guideline, source of origin
should be provided for any code or package used during the
evaluation.  For example, N4 \cite{tustison2010} is a well-known 
bias correction algorithm first made available as a tech report
in the Insight Journal \cite{tustison2009a} for subsequent introduction
into the Insight Toolkit for public dissemination.  However, since
its inclusion, different instantiations have been made available.  
Apart from the source code that accompanied the original tech
report, N4 is also available in ANTs (the only version directly maintained
by the original authors),
it is also available as a module in Slicer,%
\footnote{
http://www.slicer.org/slicerWiki/index.php/Documentation/4.2/Modules/N4ITKBiasFieldCorrection
}
a wrapper of the slicer module in nipype,%
\footnote{
http://www.mit.edu/~satra/nipype-nightly/interfaces/generated/nipype.interfaces.slicer.filtering.n4itkbiasfieldcorrection.html
}
a module in \verb#c3d#,%
\footnote{
http://www.itksnap.org8/pmwiki/pmwiki.php?n=Convert3D.Documentation
}
and as a plugin in the BRAINS suite.%
\footnote{
http://www.nitrc.org/plugins/mwiki/index.php/brains:N4ITK
}
While each version is dependent on the original source
code, there are subtle variations which can affect performance.
For example, the parameters governing the spline distance and 
number of iterations directly affect outcome and should be tuned
based on application.  However, as a particular case, the \verb#c3d# 
implementation hard codes
these values with no access to modify them by the user potentially
leading to suboptimal performance.

\subsection{Provide computational platform details of the evaluation}
An recent article \cite{gronenschild2012} pointed out significant
differences in Freesurfer output that varied with Freesurfer release
version and with operating systems.  While the former is to be
expected given upgrades and big fixes which occur between releases,
the latter underscores both the need for consistency in study processing
as well as the reporting of computational details for reproducibility.

\subsection{Pre- and post-processing steps need to be supplied.}
In addition to disclosure of all parameters associated with the
methods to be compared, all processing steps from the raw 
to the final processed images in the workflow
need to be specified.    Justification for any deviation
of steps between algorithms needs to be provided.  Additionally,
failures during processing should be mentioned and whether
or not these data were excluded from subsequent analysis.

\subsection{Comparisons should be performed on publicly available data.}
For reasons of reproducibility and transparency, evaluations should be performed
using publicly availabile data sets.  Given the rather large number
of such institutional efforts including 
IXI,%
\footnote{
http://www.brain-development.org
}
NKI,%
\footnote{
http://fcon\_1000.projects.nitrc.org/indi/pro/nki.html
}
OASIS,%
\footnote{
http://www.oasis-brains.org
}
Kirby,%
\footnote{
http://mri.kennedykrieger.org/databases.html
},
LONI,%
\footnote{
http://www.loni.ucla.edu/Research/Databases/
}, and
others, evaluations should include (if not be exhausted by)
comparisons using such data.  While evaluation on specific abnormal cohorts
might exclude such possibilities, such specific evaluations should
be extensively motivated in the introduction and/or discussion.
For example, if a particular algorithm with general application is 
found to perform better
on a private cohort of Parkinson's disease subject data, reasons
for performance disparity should be offered and supplemented with
analysis on public data.

\subsection{Resulting images should be publicly posted}

@@

\subsection{Properly contextualize comparisons and motivate performance differences.}
In addition to algorithmic and study specifics, it is important to 
discuss potential limitations concerning qualitative and/or quantitative 
assessment metrics.  In \cite{rohlfing2012}, Rohlfing pointed out 
deficiencies in using standard overlap measures and image similarity
metrics in quantifying performance.  Other issues, such as biological
plausibility of the resulting transforms, need to also be considered.

@@

\subsection{Co-authors should verify findings.}

>>>>

\section{Guidelines}
A solid evaluation is essential to a paper's longevity and
impact.  This requires effort in both planning the evaluation (before any
experiment is performed) and in publishing the evaluation.
We suggest general guidelines for both of these steps most of which
derive from basic scientific principles of clarity and reproducibility.

\subsection{Designing the Evaluation Study}

\subsubsection{Define the Algorithm's Context}
The only algorithm that is capable of outperforming all others on all
problems is exhaustive search.  Therefore, many algorithms employ
prior knowledge such as regularization or other heuristics that
restrict the search space.  The very idea that one, for instance,
registration algorithm should be better than all other algorithms on
all types of data is fundamentally flawed.  Indeed, the No Free Lunch
Theorems provide bounds on solution quality that show otherwise
i.e. specifically that ``improvement of performance in problem-solving hinges on using prior information to match procedures to problems.''  One should keep this in
mind when following the next guideline.

\subsubsection{Hassle the developers for a set of default parameters.}
Dovetailing with the previous item, developers should also facilitate
comparisons by providing default parameter files, sample scripts, 
or possible command line calls.  For example, the developers of
the image registration program Elastix \cite{klein2010} provide an
assortment of parameter files on the wiki page%
\footnote{
  http://elastix.bigr.nl/wiki/index.php/Parameter\_file\_database
}
listed in tabular format complete with  short
description (including applied modality and object of interest) and 
any publications which used that specific parameter file.  Additionally,
other packages are associated with discussion forums where users can
query regarding usage.

\subsubsection{Do not code up your own version of the algorithm particularly 
if one is available from the original authors.  If you do, be sure to make
your implementation available.}

\subsubsection{Co-authors should verify findings.}

\subsubsection{Comparisons should be performed on publicly available data.}
For reasons of reproducibility and transparency, evaluations should be performed
using publicly availabile data sets.  Given the rather large number
of such institutional efforts including 
IXI,%
\footnote{
http://www.brain-development.org
}
NKI,%
\footnote{
http://fcon\_1000.projects.nitrc.org/indi/pro/nki.html
}
OASIS,%
\footnote{
http://www.oasis-brains.org
}
Kirby,%
\footnote{
http://mri.kennedykrieger.org/databases.html
},
LONI,%
\footnote{
http://www.loni.ucla.edu/Research/Databases/
}, and
others, evaluations should include (if not be exhausted by)
comparisons using such data.  While evaluation on specific abnormal cohorts
might exclude such possibilities, such specific evaluations should
be extensively motivated in the introduction and/or discussion.
For example, if a particular algorithm with general application is 
found to perform better
on a private cohort of Parkinson's disease subject data, reasons
for performance disparity should be offered and supplemented with
analysis on public data.

\subsection{Publishing the Evaluation}

\subsubsection{Include parameters.}
A best practice exemplar of comparative evaluation studies
is the work by Klein et al. \cite{Klein2009} in which
he compared 14 nonlinear registration algorithms using
four publicly available labeled brain data sets.  
As part of the study, the authors themselves tuned 
the parameters for their algorithm to ensure good
performance.  As
part of this work, Prof. Klein availed the scripts
used for the study on his website.%
\footnote{
http://mindboggle.info/papers/evaluation\_NeuroImage2009.php
}
This practice promotes reproducibility, fairness, and transparency in 
the evaluation and is perhaps most fundamental in these types
of studies for establishing conclusions regarding the relative
performance of algorithms.

@@

\subsubsection{Provide details as to the source of the algorithm}
As a corollary to the previous guideline, source of origin
should be provided for any code or package used during the
evaluation.  For example, N4 \cite{tustison2010} is a well-known 
bias correction algorithm first made available as a tech report
in the Insight Journal \cite{tustison2009a} for subsequent introduction
into the Insight Toolkit for public dissemination.  However, since
its inclusion, different instantiations have been made available.  
Apart from the source code that accompanied the original tech
report, N4 is also available in ANTs (the only version directly maintained
by the original authors),
it is also available as a module in Slicer,%
\footnote{
http://www.slicer.org/slicerWiki/index.php/Documentation/4.2/Modules/N4ITKBiasFieldCorrection
}
a wrapper of the slicer module in nipype,%
\footnote{
http://www.mit.edu/~satra/nipype-nightly/interfaces/generated/nipype.interfaces.slicer.filtering.n4itkbiasfieldcorrection.html
}
a module in \verb#c3d#,%
\footnote{
http://www.itksnap.org8/pmwiki/pmwiki.php?n=Convert3D.Documentation
}
and as a plugin in the BRAINS suite.%
\footnote{
http://www.nitrc.org/plugins/mwiki/index.php/brains:N4ITK
}
While each version is dependent on the original source
code, there are subtle variations which can affect performance.
For example, the parameters governing the spline distance and 
number of iterations directly affect outcome and should be tuned
based on application.  However, as a particular case, the \verb#c3d# 
implementation hard codes
these values with no access to modify them by the user potentially
leading to suboptimal performance.

\subsubsection{Provide computational platform details of the evaluation}
An recent article \cite{gronenschild2012} pointed out significant
differences in Freesurfer output that varied with Freesurfer release
version and with operating systems.  While the former is to be
expected given upgrades and big fixes which occur between releases,
the latter underscores both the need for consistency in study processing
as well as the reporting of computational details for reproducibility.

\subsubsection{Pre- and post-processing steps need to be supplied.}
In addition to disclosure of all parameters associated with the
methods to be compared, all processing steps from the raw 
to the final processed images in the workflow
need to be specified.    Justification for any deviation
of steps between algorithms needs to be provided.  Additionally,
failures during processing should be mentioned and whether
or not these data were excluded from subsequent analysis.

\subsubsection{Resulting images should be publicly posted}

@@

figshare

\subsubsection{Properly contextualize comparisons and motivate performance differences.}
In addition to algorithmic and study specifics, it is important to 
discuss potential limitations concerning qualitative and/or quantitative 
assessment metrics.  In \cite{rohlfing2012}, Rohlfing pointed out 
deficiencies in using standard overlap measures and image similarity
metrics in quantifying performance.  Other issues, such as biological
plausibility of the resulting transforms, need to also be considered.

@@

\subsection{Post-Publication: Publications Age}

quoting satra ``a published evaluation is old the day the evaluation was done, as most software keeps changing. for example, arno's results were fine at that time with those versions and parameterization of the software, but might be a different story currently.''
